---
title: "TED Talks by Phil Hansen and Vijay Kumar."
author: "Agunbiade, Adedoyin"
subtitle: "Sentiment analysis using R-programming."
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
### Don't delete this setup code chunk from your file
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NULL)   ## DON'T ALTER THIS: this is to prevent printing the code or any unnecessary addition in your final "html" report file.

# You can extend this list below to load all the packages required for your analyses:
#====================================================================================
library(knitr)
library(dsEssex)
library(tidyverse)
library(tidytext)
library(ggrepel)


# load the 'ted_talks' data
#=========================
data(ted_talks)
```

```{=html}
<style>
body {
text-align:justify;
}
</style>
```
## Introduction
TED talks are influential public speeches delivered by experts from various fields. Phil Hansen, a multimedia artist, delivered a TED talk in 2013 titled “Embrace the Shake” where he discussed how he transformed his limitations into new possibilities by being creative, and he encourages others to do the same. His talk was inspired by his diagnosis of permanent nerve damage that prevented him from doing art. Vijay Kumar, a roboticist, delivered a talk in 2012 titled “Robots that Fly… and Cooperate.” He discussed the challenges of building agile robots, the technologies involved, and their potential to revolutionise society. His robots can be used as first respondents or in construction, and their diversity and agility make them stand out. This report compares and analyses Phil Hansen and Vijay Kumar’s topics, ideas, and contexts using text analysis tools in R programming.

## Methods
***Install Packages and Load Data:*** This analysis used R packages including `tidyverse`, `dsEssex`, `tidytext`, `ggrepel`, and `knitr`. The `ted_talk` dataset was loaded and filtered using the `filter()` function to include only talks by Phil Hansen and Vijay Kumar.

***Pre-processing:*** To ensure consistency, the transcript data was cleaned by converting all characters to lowercase, removing unwanted characters, whitespaces, punctuations, and numbers. The unnest_tokens() function was used to tokenise the transcripts into individual words and convert data to tidy format thus enabling analysis with the packages. Stop-words were then removed using the `anti_join()` function on get_stopwords() to improve analysis efficiency and focus on meaningful words.

***Word Analysis:*** Post-preprocessing, `tidyverse` package functions were used to determine word frequency and common words. Sentiment analysis was done via the `bing` and `nrc` lexicons to identify emotions in each speaker's speech. Log-odd ratios were calculated for each speaker's words using a custom compute_OR function, which includes a smoothing factor of 0.5 to address zero proportions. If the odds ratio >1, the first speaker is more likely to express the sentiment, while an odds ratio <1 suggests the second speaker is more likely to express the sentiment.

***Data visualisation:*** `ggplot2` in `tidyverse` and `Knitr` were used to create barplot, scatterplot, and tables to help get better data insights and interpretation.

```{r, fig.height=4, fig.width=7}

# Filter data for Phil Hansen and Vijay Kumar
MyData <- ted_talks %>%
  filter(speaker %in% c("Phil Hansen", "Vijay Kumar"))

#clean the data 
MyData$text <- str_replace_all(MyData$text, "[^[:alpha:][:space:]]", "")  # remove unwanted characters such as numbers, and punctuations
MyData$text<- (MyData$text = tolower(MyData$text)) # convert text to lowercase

#remove stopwords from data
tidyData <- MyData %>%
  unnest_tokens(word, text) %>% # unnest_tokens() converts the text to individual words.
  anti_join(get_stopwords())  # anti_join() used on get_stopwords() helps to remove the stop words
```

## Result

***Word Frequency analysis.***

The initial word count between the speakers was 3886, which was reduced to 1877 after removing stop-words, indicating almost half of the words were irrelevant. Vijay Kumar used more words (1105) than Phil Hansen (772), suggesting Vijay Kumar used a greater number of vocabularies. This may be due to his longer talk duration and since this data did not consider the speech length, this result may be biased.

```{r, fig.height=4, fig.width=7}
#total words before the removal of stop words
total_words_b4_stop <- MyData %>%
  unnest_tokens(word, text) %>%
  summarize(total_words_b4_stop = n()) #here summarise is used to give a summary of the total word

#total words after the removal of stop words
total_word_after_stop<- MyData %>%
  unnest_tokens(word, text) %>%
  anti_join(get_stopwords()) %>%
  summarize(total_words_b4_stop = n()) 

# Data after removal of stop words 
tidyData <- MyData %>%
  unnest_tokens(word, text) %>%
  anti_join(get_stopwords()) 
```

Figure 1 below shows the most common words used in Phil Hansen's and Vijay Kumar's TED talks.

```{r, fig.height=4, fig.width=7}
#Count the frequency of words between the two speakers
speakers_top <- tidyData %>%
  group_by(speaker)%>% #groups the data by the attribute "speaker"
  count(word, sort = TRUE) %>% #counts the frequency of words and sort in descending order
  slice_max(n, n=10) # View the top 10 words for each speaker
  
plot_1<-tidyData %>%
  group_by(speaker)%>% 
  count(word, sort = TRUE) %>% 
  slice_max(n, n=10) %>%
  #visualise the top words using ggplot
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col() + 
  labs(title = "Figure 1: Most Frequent Words for both speakers",
                    x = "Frequency",
                    y = "Words") #customises the axis and title
plot_1
```

Figure 2 below also displays the word frequency analysis. The result shows that Vijay Kumar used the words "robot" (39 times), "robots" (34 times), and "can" (22 times) more frequently, while Phil Hansen used "art" (25 times), "shake" (13 times), and "just" (12 times) more often. The words "go" and "actually" were common between the two speakers, with "actually" being used more by Phil Hansen (8 times versus 7 times) and "go" was used more by Vijay Kumar (6 times versus 5 times). These findings indicate that both speakers have different approaches and perspectives on their topic, with Phil Hansen focusing on art and Vijay Kumar on robotics, and that their most frequently used words were related to their respective fields.

```{r, fig.height=4, fig.width=7}
#compare common words and visualise
compare_talk<- tidyData %>%
  count(speaker, word)%>%   
  group_by(word) %>%                                             
  filter(sum(n) > 10) %>% #filters for words that is more than 10
  ungroup() %>%                                             
  pivot_wider(names_from = "speaker", values_from = "n", values_fill = 0) #converts data to wide format

#create a scatterplot to visualise result using ggplot available in tidyverse
plot_2 <- compare_talk %>% ggplot(aes(`Phil Hansen`, `Vijay Kumar`)) +
  geom_abline(color = "red", size = 1.2, alpha = 0.8, lty = 2) +
  geom_text_repel(aes(label = word), max.overlaps = 15) +
  labs(title = "Figure 2: Comparison of words between Phil Hansen and Vijay Kumar") +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, 4)) #customise scale

plot_2
```

***Sentiment analysis and log-odd ratio***

Using the `bing` lexicon, it was displayed in Figure 3 that both speakers used both positive and negative sentiments. Some of the most commonly used negative sentiments were “shake” and “destruction while the most positively used sentiments were “like” and autonomous. Despite both speakers using positive and negative sentiments, their respective frequencies varied. Table 1 for instance demonstrates that Phil Hansen had a higher percentage of negative sentiments (63.8%) than positive sentiments (36.2%), while Vijay Kumar had a higher percentage of positive sentiments (68.7%) than negative sentiments (31.2%). When the speaker's sentiments were compared to each other, Vijay Kumar's talk had more positive sentiments (57) than Phil Hansen's talk (34), while Phil Hansen had a higher negative sentiment count (60) compared to Vijay Kumar's (26). Phil Hansen used words such as "creative" and "best" in a positive sense and words such as "shake" and "destruction" in a negative connotation. In contrast, Vijay Kumar used words such as "like" and "autonomous" positively and words such as "object" and "complicated" in a negative tone. Overall, these differences suggest that the speakers had different approaches to conveying their messages to the audience, which could influence how the audience perceives them.

```{r, fig.height=4, fig.width=7}
#compare speakers sentiments using bing lexicon from the tidytext package
speakers_sentiments_BING <- 
  tidyData %>%
  inner_join(get_sentiments("bing"), by = "word")%>%
  count(speaker, sentiment) %>%
  pivot_wider(names_from = speaker, values_from = n, values_fill = 0)
  
#calculate log-odd ratio for the bing lexion to identify association of sentiments
log_sentiment_bing<- speakers_sentiments_BING %>% 
  mutate(OR = compute_OR(`Phil Hansen`, `Vijay Kumar`, correction = FALSE), #uses compute_OR function created
         log_OR = log(OR), sentiment = reorder(sentiment, log_OR)) %>%
  select(sentiment, `Phil Hansen`, `Vijay Kumar`, OR, log_OR) %>% 
  #visualise result in a table using Knit package available in R 
  kable(caption = "Table 1: Sentiment analysis and log-odd ratio of Phil Hansen and Vijay Kumar TED talk using BING lexicon") # kable () is the part of the knitr package

log_sentiment_bing
```

```{r, fig.height=4, fig.width=7}
#identify most common sentiment for each speaker
top_sentiment_BING<- tidyData %>%
  inner_join(get_sentiments("bing"))%>% # the inner join function merges the sentiments
  count(word, speaker,sentiment) %>%
  # group by sentiment
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup() # result depicting each speaker's top sentiment was not printed in the report. This code is to help identify them.


#visualise top negative and positive sentiment for BOTH speakers
Plot_3<- top_sentiment_BING %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Figure 3: Common sentiments for both speakers using bing lexicon")+
  facet_wrap(~ sentiment, scales = "free")# puts the plot side by side
  
Plot_3
#this plot shows sentiment for both speakers
```

The log-odd ratio used to measure the odds of the sentiments conveyed helped in understanding the strength of association between a particular sentiment. The odd ratio (OR) and the log_odd ratio (log_OR) depicted in Table 1 above, support the analysis earlier described. The OR and log_OR for negative sentiment between Phil Hansen and Vijay Kumar were 3.86 and 1.35, respectively. This suggests that Phil Hansen is more likely to express negative emotions while Vijay Kumar is less likely to. Contrarily, the OR and log_OR for positive sentiments between the speakers were 0.25 and -1.35, respectively. Overall, the bing lexicon identifies Vijay Kumar as having a higher probability of expressing positive emotions, thus suggesting he is more optimistic, while Phil Hansen's words were more likely to express negative emotions suggesting his speech is more critical.

Furthermore, by using the `nrc` lexicon to identify the emotions conveyed, it was observed that the speakers used a range of sentiments, indicating that their talks were not entirely positive or negative. Table 2 result displays a slightly different result from Table 1's result. Analysing Table 2 and Figure 4 shows that even though Phil Hansen had a higher negative and positive sentiment count, these emotions are more associated with Vijay Kumar as the ORs were less than 1 and the log_ORs were negative. Table 2 result identifies that both speakers expressed anger and anticipation, but Phil Hansen exhibits these sentiments more. Also, the results demonstrate that both speakers equally expressed disgust (sentiment count of 8); however, this sentiment is more associated with Vijay Kumar, as shown in Figure 4. Figure 4 also shows that the most associated emotion for Vijay Kumar is "trust" while "surprise" was more associated with Phil Hansen.

The two lexicons displaying slightly different analysis highlights the limitations of lexicons as its subjective and systematically different in word searches. Similarly, emotions conveyed may vary in a context that may not be reflected by lexicons. For instance, the word "trick" was used by Vijay positively; however, this was analysed as negative by the `bing` sentiment.

Another limitation of this analysis was that it was based solely on text and did not factor speaker's tone, body language, and visuals which could give better insight into the speaker's emotions and topic.

```{r, fig.height=4, fig.width=7}
#defining a function to calculate log-odd ratio
compute_OR <- function(num, denom, correction = TRUE){
  if(correction){
    ODDS_N = (num + 0.5) / (sum(num) - num + 0.5) #0.5 added for smoothing
    ODDS_D = (denom + 0.5) / (sum(denom) - denom + 0.5)
    } else {
      ODDS_N = num / (sum(num) - num)
      ODDS_D = denom / (sum(denom) - denom)
      }
  OR = ODDS_N/ODDS_D}
```

```{r, fig.height=4, fig.width=7}

#compare sentiment analysis for both speakers using nrc lexicon
compare_sentiment_NRC<-tidyData %>%
  inner_join(get_sentiments(lexicon="nrc"), by = "word")%>% #nrc lexicon used to identify other emotions
  count(speaker, sentiment) %>%
  pivot_wider(names_from = speaker, values_from = n, values_fill = 0)%>%
  #calulate log odd ratio for the nrc lexicon
  mutate(OR = compute_OR(`Phil Hansen`, `Vijay Kumar`, correction = FALSE), #uses compute_OR function created
         log_OR = log(OR), sentiment = reorder(sentiment, log_OR)) %>%
  select(sentiment, `Phil Hansen`, `Vijay Kumar`, OR, log_OR) %>% 
  #visualise result in a table
  kable(caption = "Table 2: Common sentiments and log-odd ratio between Phil Hansen and Vijay Kumar using nrc") 
 
compare_sentiment_NRC


#visualise the association and log-odd ratio using ggplot
plot_4<- tidyData %>%
  inner_join(get_sentiments("nrc"), by = "word")%>%
  count(speaker, sentiment) %>%
  pivot_wider(names_from = speaker, values_from = n, values_fill = 0) %>%
  mutate(OR = compute_OR(`Phil Hansen`, `Vijay Kumar`, correction = FALSE), 
         log_OR = log(OR), sentiment = reorder(sentiment, log_OR)) %>%
  ggplot(aes(sentiment, log_OR, fill = log_OR < 0)) +
  geom_col(show.legend = FALSE) +
  ylab("Log odds ratio") + ggtitle("Figure 4: The association between sentiments using nrc lexicon") +
  coord_flip() + 
  scale_fill_manual(name = "", values = c("darkblue", "darkred"))

plot_4
```

## Conclusion

Overall, this text analysis helped gain insight into topics, ideas, context, and emotions involved in Phil Hansen's and Vijay Kumar's TED talk. Both speakers had different perspectives and unique speaking styles, and their words related to their fields. Different emotions were used to convey their message; however, more analysis and other factors need to be considered to gain a better understanding.


